---
title: "Darwin Core mapping"
subtitle: "For: Bat Data Bulgaria"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(lubridate)      # to transform the date
library(mapview)
```

# Read source data

Create a data frame `input_data` from the source data:

```{r}
input_data <- read_csv(here::here("datasets", "batData", "data", "raw", "bat-data-Bg-NT-20250207Refine.csv"))
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```



## Scientific names (not applicable)

Add columns:

```{r}
input_data %<>%
  mutate(dwc_kingdom = "Animalia") 
```

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:


```{r}
parsed_names <- input_data %>%
  distinct(`species`) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

Correct names and reparse:

##all names are correct


## Taxon ranks (not applicable)

The nameparser function also provides information about the rank of the taxon (in `rankmarker`). Here we join this information with our checklist. Cleaning these ranks will done in the Taxon Core mapping:

```{r}
input_data %<>% left_join(
  parsed_names %>%
  select(scientificname, rankmarker),
  by = c("species" = "scientificname"))
```

```{r}
# Adjust rankmarker
input_data %<>% mutate(rankmarker = recode(species,
  "Myotis myotis/blythii" = "genus",
  "Myotis sp." = "genus",
  "Rhinolophus sp." = "genus",
  "Pipistrellus sp." = "genus",
  "Plecotus sp." = "genus",
  "P.kuhlii/P.nathusii" = "genus",
  .default = rankmarker
))
```

## Taxon IDs (not applicable)



## Preview data

Show the number of taxa and distributions per kingdom and rank:

```{r}
#input_data %>%
#  group_by(kingdom, rankmarker) %>%
#  summarize(
#    `# taxa` = n_distinct(taxon_id),
#    `# distributions` = n()
#  ) %>%
#  adorn_totals("row")
```

Preview data:

```{r}
input_data %>% head()
```

# Darwin Core mapping

## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Bulgarian Bat data")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "placeholder") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "Bulgarian Bat Society")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Bat data of Bulgaria")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```


```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_basisOfRecord = case_when(sourceEn == 'Literature data' ~ "materialSample",
                                                      .default = as.character("humanObservation")))
```

The following terms contain information about the taxon:

### occurrenceID


```{r}
occurrence %<>% mutate(dwc_occurrenceID = ID)
```


```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_occurrenceID = paste0("BUL:BAT:OBS:", row_number()))
```

### associatedReferences

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_associatedReferences = notes)
```


### recordedBy

this piece will also remove the "|"

```{r eval=FALSE, include=FALSE}

occurrence %<>% mutate(
  dwc_recordedBy = apply(select(., `data owner 1`, `data owner 2`, `data owner 3`), 1, function(x) paste(discard(x, is.na), collapse = " | "))
)

```


```{r}
occurrence %<>% mutate(
  dwc_recordedBy = paste0(
    coalesce(`data owner 1`, ""), " | ",
    coalesce(`data owner 2`, ""), " | ",
    coalesce(`data owner 3`, "")
  )
)
```




```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_recordedBy = paste0(`data owner 1`, " | ", `data owner 2`, " | ", `data owner 3`)) %>% 
``` 




### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = counts)
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = case_when(counts > 0 ~"present",
                                                        counts == 0 ~"absent"))
```



### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = eventDate)
  
```



### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = method)
```



### samplingEffort


 

### locationID


### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BG")
```

### stateProvince

### locality

```{r}
occurrence %<>% mutate(dwc_locality = `site name`)
```


### Municipality

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_municipality = location )
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = decimalLatitude) 
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = decimalLongitude ) 
```

### altitude

```{r}
occurrence %<>% mutate(dwc_minimumElevationInMeters = `alt (m)` ) 
```

```{r}
occurrence %<>% mutate(dwc_maximumElevationInMeters = `alt (m)` ) 
```


```{r}
occurrence %<>% mutate(dwc_verbatimElevation = `alt (m)` ) 
```

### locationRemarks

```{r}
occurrence %<>% mutate(dwc_locationRemarks = `site type` ) 
```



### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### georeferenceRemarks

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_georeferenceRemarks = "points are centroid of 10km UTM grid")
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```



### scientificName & previous identifications

```{r}
occurrence %<>% mutate(dwc_previousIdentifications = species)
```



```{r}
# Adjust scientificName
input_data %<>% mutate(scientificName = recode(species,
  "Myotis myotis/blythii" = "Myotis",
  "Myotis sp." = "Myotis",
  "Rhinolophus sp." = "Rhinolophus",
  "Pipistrellus sp." = "Pipistrellus",
  "Plecotus sp." = "Plecotus",
  "P.kuhlii/P.nathusii" = "Pipistrellus",
  .default = species
))
```





### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = rankmarker)

```

### scientificNameAuthorship




### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```




## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

MAke sure records are unique

```{r}
occurrence <- unique(occurrence)
```

```{r eval=FALSE, include=FALSE}
occurrence <- occurrence %<>%
  mutate(across(everything(), ~ ifelse(is.na(.), , .)))
```


## this is changing Df to something we dont need
```{r eval=FALSE, include=FALSE}
occurrence <- lapply(occurrence, as.character) 
```



```{r}
occurrence <- occurrence %<>%
  mutate(across(where(is.factor), as.character))
```

```{r eval=FALSE, include=FALSE}
occurrence <- occurrence %<>%
  mutate(across(everything(), ~ ifelse(is.na(.), , .)))
```

```{r}
occurrence <- occurrence %>% drop_na()
```



Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "batData", "data", "processed", "occurrence.csv"), na = "")

```


# visualisation

Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

## Filter for not missing values

```{r}
occurrenceMap %<>%
        filter(decimalLongitude > 0 ) 
        
#head (occurrence, n = 5 )

```



```{r}
head(occurrenceMap)
```


```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```