---
title: "Darwin Core mapping"
subtitle: "For: HabiProt Counting"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(lubridate)      # to transform the date
library (sf)
library(mapview)
library(mgrs)
library(stringr)
```

# Read source data

Create a data frame `input_data` from the source data:

```{r eval=FALSE, include=FALSE}
input_data <- read_excel(here::here("datasets", "habiprot", "data", "raw", "HabiProtData.xlsx"), guess_max = 400000)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```


# Darwin Core mapping

## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by-nc/4.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "HabiProt")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "placeholder") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "HabiProt")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Artropod observations derived from HabiProt Alciphron database in Serbia")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

The following terms contain information about the taxon:

### occurrenceID



```{r}
occurrence %<>% mutate(dwc_occurrenceID = paste0("HABIPROT:OBS:", id))
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = author)
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = identifier)
```



### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = ifelse(as.numeric(number) == 0, NA, as.numeric(number)))
```

### eventDate

```{r}
occurrence <- occurrence %<>% 
  mutate(
    # eerst alles naar character
    date_char = as.character(date),
    # detecteer getallen (Excel serials) en zet om
    dwc_eventDate = ifelse(
      suppressWarnings(!is.na(as.numeric(date_char))),
      format(as.Date(as.numeric(date_char), origin = "1899-12-30"), "%Y-%m-%d"),
      format(parse_date_time(date_char, orders = c("Ymd", "Y-m-d", "d/m/Y", "d-m-Y")), "%Y-%m-%d")
    )
  ) %>%
  select(-date_char) # hulpkolom weg
```

### lifeStage

```{r}
occurrence %<>% mutate(
    dwc_lifestage = case_when(
      stadium == "I" ~ "Imago",
      stadium == "O" ~ "ovum",
      stadium == "L" ~ "larva",
      stadium == "P" ~ "pupa",
      stadium == "M" ~ "dead specimen",
      TRUE ~ NA_character_  # alles wat niet in de lijst staat wordt NA
    )
  )
```


### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "casual observation")
```

### samplingEffort
 

### locationID


### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "RS")
```

### continent

```{r}
occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### stateProvince


### Municipality


### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude_verb = latitude) 
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude_verb = longitude) 
```




#transformcoordinates again

wrong calculation... it is providing eastings


```{r eval=FALSE, include=FALSE}
occurrence %<>%
  mutate(
    utm = as.character(utm),               # your short code column
    mgrs_full = paste0("34T", utm),       # full 8-char MGRS code
    latlng = lapply(mgrs_full, mgrs_to_latlng),   # convert to lat/lon
    dwc_decimalLatitude  = sapply(latlng, `[[`, 2),   # extract lat
    dwc_decimalLongitude = sapply(latlng, `[[`, 3)    # extract lon
  ) 
 # select(-mgrs_full, -latlng)
```


```{r}
occurrence %<>%
  mutate(
    utm       = as.character(utm),
    mgrs_full = paste0("34T", utm),

    # Safe wrapper for mgrs_to_latlng
    latlng_sw = lapply(mgrs_full, function(m) {
      tryCatch(mgrs_to_latlng(m), error = function(e) c(NA, NA, NA))
    }),

    sw_lat = sapply(latlng_sw, `[[`, 2),
    sw_lon = sapply(latlng_sw, `[[`, 3)
  ) %>%
  # drop rows with NA coords before sf step
  filter(!is.na(sw_lat) & !is.na(sw_lon)) %>%
  mutate(
    geom_wgs   = st_as_sf(data.frame(lon = sw_lon, lat = sw_lat),
                          coords = c("lon","lat"), crs = 4326),
    geom_utm   = st_transform(geom_wgs, 32634),
    geom_utm_c = st_geometry(geom_utm) + c(5000, 5000),
    geom_wgs_c = st_transform(st_sfc(geom_utm_c, crs = 32634), 4326),
    coords     = st_coordinates(geom_wgs_c),
    dwc_decimalLongitude = coords[,1],
    dwc_decimalLatitude  = coords[,2]
  ) %>%
  select(-mgrs_full, -latlng_sw, -sw_lat, -sw_lon,
         -geom_wgs, -geom_utm, -geom_utm_c, -geom_wgs_c, -coords)
```



### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### verbatimcoordinatesystem

```{r}
occurrence %<>% mutate(dwc_verbatimCoordinateSystem = "UTM")
```

### verbatimCoordinates

```{r}
occurrence %<>% mutate(dwc_verbatimCoordinates = utm)
```


### elevation

```{r}
occurrence %<>% mutate(dwc_minimumElevationInMeters = altitude )
```


### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "7071")
```

### georeferenceRemarks

```{r include=FALSE}
occurrence %<>% mutate(dwc_georeferenceRemarks = "coordinates are centroid of UTM10km square"
  )
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```



### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = species)
```



### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")

```

### scientificNameAuthorship


### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```





## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "habiprot", "data", "processed", "occurrence.csv"), na = "")

```

Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap <- occurrenceMap %>%
  filter(!is.na(decimalLatitude) & !is.na(decimalLongitude))
```

```{r}
occurrenceMap <- occurrenceMap %>%
  group_by(decimalLongitude, decimalLatitude) %>%
  summarise(occurrences = n(), .groups = "drop") %>%
  select(decimalLongitude, decimalLatitude, occurrences)
```


```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```


```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

```{r}
occurrenceMap2 <- occurrence
```

```{r}
occurrenceMap2 <- occurrenceMap2 %>%
  filter(!is.na(decimalLatitude_verb) & !is.na(decimalLongitude_verb))
```

```{r}
occurrenceMap2 <- occurrenceMap2 %>%
  group_by(decimalLongitude_verb, decimalLatitude_verb) %>%
  summarise(occurrences = n(), .groups = "drop") %>%
  select(decimalLongitude_verb, decimalLatitude_verb, occurrences)
```


```{r}
occurrenceMap2$decimalLatitude_verb <- as.numeric(occurrenceMap2$decimalLatitude_verb)
```


```{r}
occurrenceMap2$decimalLongitude_verb <- as.numeric(occurrenceMap2$decimalLongitude_verb)
```

```{r}
head(occurrenceMap2)
```
```{r}
mapview(occurrenceMap2, xcol = "decimalLongitude_verb", ycol = "decimalLatitude_verb", crs = 4269, grid = FALSE)

```

