---
title: "Darwin Core mapping"
subtitle: "For: HabiProt Counting"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(lubridate)      # to transform the date
```

# Read source data

Create a data frame `input_data` from the source data:

```{r include=FALSE}
input_data <- read_excel(here::here("datasets", "habiprot", "data", "raw", "HabiProtData.xlsx"), guess_max = 400000)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```


# Darwin Core mapping

## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by-nc/4.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "HabiProt")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "placeholder") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "HabiProt")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Artropod observations derived from HabiProt Alciphron database in Serbia")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

The following terms contain information about the taxon:

### occurrenceID



```{r}
occurrence %<>% mutate(dwc_occurrenceID = paste0("HABIPROT:OBS:", id))
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = author)
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = ifelse(as.numeric(number) == 0, NA, as.numeric(number)))
```

### eventDate

```{r}
occurrence <- occurrence %<>% 
  mutate(
    # eerst alles naar character
    date_char = as.character(date),
    # detecteer getallen (Excel serials) en zet om
    dwc_eventDate = ifelse(
      suppressWarnings(!is.na(as.numeric(date_char))),
      format(as.Date(as.numeric(date_char), origin = "1899-12-30"), "%Y-%m-%d"),
      format(parse_date_time(date_char, orders = c("Ymd", "Y-m-d", "d/m/Y", "d-m-Y")), "%Y-%m-%d")
    )
  ) %>%
  select(-date_char) # hulpkolom weg
```



### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "casual observation")
```

### samplingEffort
 
```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_samplingEffort = case_when(
  !is.na(Mandagen) ~ paste("daysOfWork:", round(Mandagen, digits = 2)),
  !is.na(Val_aantal) ~ paste("numberOfTraps:", floor(Val_aantal)),
  TRUE ~ ""
))
```

### locationID


### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "RS")
```

### stateProvince


### Municipality


### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = latitude) 
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = longitude) 
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "7071")
```

### georeferenceRemarks

```{r include=FALSE}
occurrence %<>% mutate(dwc_georeferenceRemarks = "coordinates are centroid of UTM10km square"
  )
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```



### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = species)
```



### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")

```

### scientificNameAuthorship


### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```






## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "albania", "data", "processed", "occurrence.csv"), na = "")

```
