---
title: "Darwin Core Mapping"
subtitle: "THF - Mammals Montenegro"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv


```{r}
csv_files <- list.files(here::here("datasets", "Mammals_montenegro", "data", "raw"), pattern = "*.csv$", full.names = TRUE)

```

```{r}
data <- lapply(csv_files, read_csv) %>% bind_rows()
```


```{r}
#data <- bind_rows(csv_files)
```

```{r}
input_interim <- data
```


Preview data:
```{r}
input_interim %>% head(n = 5)
```


# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```


# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/XXXXX")
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Wildlife Montenegro") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "Wildlife Montenegro") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Wildlife Montenegro, mammals in Montenegro") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = case_when (Metoda == 'Camera trap' ~ "machine observation",
                                                  .default = "humanObservation" ))
```

```{r}
occurrence %<>% mutate(dwc_occurrenceID = `__id`)

```

# recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = `Name and surname`)

```

# individualquantity

```{r}
occurrence %<>% mutate(dwc_organismQuantity = case_when(brojnost > 0 ~ brojnost,
                                                        is.na(brojnost) ~ ">= 1"))

```

# individualQuantityType

```{r}
occurrence %<>% mutate(dwc_organismQuantityType = "count - countRange")

```



```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_organismQuantity > 0 ~ "Present",
                                        dwc_organismQuantity == 0 ~ "Absent",
                                        dwc_organismQuantity == ">= 1" ~"present",
                                        is.na(dwc_organismQuantity) ~ "Present" ))

```

### informationWithHeld

### dataGeneralizations

### organismQuantity

### sex

### lifeStage

### behavior

### occurrenceRemarks

```{r}
occurrence %<>% mutate(dwc_occurrenceRemarks = occurrenceRemarks)
```

## samplingProtocol

```{r}
occurrence %<>% mutate(samplingEffort = Metoda)
```


## eventDate


```{r}
occurrence %<>% mutate(dwc_eventDate = eventDate )

```
## eventTime

```{r}
occurrence %<>% mutate(dwc_eventTime = time )

```
## habiat

```{r}
occurrence %<>% mutate(dwc_habitat = `tip habitata` )

```


# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = decimalLongitude) %>%
  mutate(dwc_decimalLatitude = decimalLatitude) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "30") %>%
  mutate(countryCode = "YI")  %>%            
  mutate(continent = "Europe")
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```


### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = `Name and surname`)
```

### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```

## verbatim identification

```{r}
occurrence %<>% mutate(dwc_verbatimIdentification = `fonetska grupa`)
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = taxonRank)
```
## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns


Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_delim(occurrence, here::here("datasets", "Mammals_montenegro", "data", "processed", "occurrence.csv"),delim = ";", na = "")
```


Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

