---
title: "Darwin Core mapping"
subtitle: "For: datozoka Counting"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(lubridate)      # to transform the date
library(sf)
library(mapview)
library(mgrs)
library(stringr)
library(uuid)
```

# Read source data

Create a data frame `input_data` from the source data:

```{r eval=FALSE, include=FALSE}
input_data <- read_excel(here::here("datasets", "DAtaZokaBulgaria", "data", "raw", "DataZoka2013Trigrad_Bulgaria.xlsx"), guess_max = 400000)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```


# filter for individualCount > 0

```{r}
input_data <- input_data %>%
  filter(`N ind` > 0)
```


# Darwin Core mapping

```{r}
input_data$occurrenceID <- sapply(1:nrow(input_data), function(x) UUIDgenerate())
```


## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by-nc/4.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Dutch Mammal society")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/w5rg22") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode


### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Dutch Mammal society summercamp records: Bulgaria")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

The following terms contain information about the taxon:

### occurrenceID



```{r}
occurrence %<>% mutate(dwc_occurrenceID = occurrenceID)
```

### recordedBy

```{r}
occurrence %<>%
  unite("recordedBy", `Observer 1`:`Observer 5`, na.rm = TRUE, sep = "; ")
```




### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = ifelse(as.numeric(`N ind`) == 0, NA, as.numeric(`N ind`)))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = Date)
```


```{r}
dms_to_decimal <- function(dms) {
  # Extract numbers
  parts <- str_match(dms, "(\\d+)°(\\d+)'(\\d+(?:\\.\\d+)?)\"?([NSEO])")
  degrees <- as.numeric(parts[,2])
  minutes <- as.numeric(parts[,3])
  seconds <- as.numeric(parts[,4])
  hemi <- parts[,5]
  
  # Convert to decimal
  dec <- degrees + minutes/60 + seconds/3600
  
  # Apply sign for S or W (O in your case)
  dec[hemi %in% c("S", "O")] <- dec[hemi %in% c("S", "O")]
  
  return(dec)
}
```

```{r}
occurrence$dwc_decimalLatitude <- dms_to_decimal(occurrence$lat)
occurrence$dwc_decimalLongitude <- dms_to_decimal(occurrence$long)

```



```{r eval=FALSE, include=FALSE}
occurrence <- occurrence %<>% 
  mutate(
    # eerst alles naar character
    date_char = as.character(date),
    # detecteer getallen (Excel serials) en zet om
    dwc_eventDate = ifelse(
      suppressWarnings(!is.na(as.numeric(date_char))),
      format(as.Date(as.numeric(date_char), origin = "1899-12-30"), "%Y-%m-%d"),
      format(parse_date_time(date_char, orders = c("Ymd", "Y-m-d", "d/m/Y", "d-m-Y")), "%Y-%m-%d")
    )
  ) %>%
  select(-date_char) # hulpkolom weg
```

### lifeStage

```{r}
occurrence %<>% mutate(
    dwc_lifestage = case_when(
      stage == "adult" ~ "adult",
      stage == "juvenile" ~ "juvenile",
      stage == "subadult" ~ "subadult",
      stage == "P" ~ "pupa",
      stage == "M" ~ "dead specimen",
      TRUE ~ NA_character_  # alles wat niet in de lijst staat wordt NA
    )
  )
```


### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = Method)
```

### samplingEffort
 

### locationID


### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BG")
```

### continent

```{r}
occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### stateProvince


### Municipality


### decimalLatitude


### decimalLongitude



#transformcoordinates again



### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### verbatimcoordinatesystem


### verbatimCoordinates



### elevation



### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### georeferenceRemarks


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```



### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = Species)
```



### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")

```

### scientificNameAuthorship


### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```





## Post-processing

# create species list

```{r}
species_list <- occurrence %>%
  distinct(dwc_scientificName) %>%
  pull(dwc_scientificName)
```

```{r}
species_summary <- input_data %>%
  group_by(Species) %>%          # Group by species
  summarise(count = n(), .groups = "drop") %>%  # Count occurrences
  arrange(Species)     
```



Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "DAtaZokaBulgaria", "data", "processed", "occurrence.csv"), na = "")

```

Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap <- occurrenceMap %>%
  filter(!is.na(decimalLatitude) & !is.na(decimalLongitude))
```

```{r}
occurrenceMap <- occurrenceMap %>%
  group_by(decimalLongitude, decimalLatitude) %>%
  summarise(occurrences = n(), .groups = "drop") %>%
  select(decimalLongitude, decimalLatitude, occurrences)
```


```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```


```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

```{r}
occurrenceMap2 <- occurrence
```

```{r}
occurrenceMap2 <- occurrenceMap2 %>%
  filter(!is.na(decimalLatitude_verb) & !is.na(decimalLongitude_verb))
```

```{r}
occurrenceMap2 <- occurrenceMap2 %>%
  group_by(decimalLongitude_verb, decimalLatitude_verb) %>%
  summarise(occurrences = n(), .groups = "drop") %>%
  select(decimalLongitude_verb, decimalLatitude_verb, occurrences)
```


```{r}
occurrenceMap2$decimalLatitude_verb <- as.numeric(occurrenceMap2$decimalLatitude_verb)
```


```{r}
occurrenceMap2$decimalLongitude_verb <- as.numeric(occurrenceMap2$decimalLongitude_verb)
```

```{r}
head(occurrenceMap2)
```
```{r}
mapview(occurrenceMap2, xcol = "decimalLongitude_verb", ycol = "decimalLatitude_verb", crs = 4269, grid = FALSE)

```

