---
title: "Darwin Core mapping"
subtitle: "For: Albania Bird Counting"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(lubridate)      # to transform the date
```

# Read source data

Create a data frame `input_data` from the source data:

```{r}
input_data <- read_csv(here::here("datasets", "albania", "data", "raw", "PPNEA_ledi_birds.csv"), guess_max = 400000)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```



## Scientific names (not applicable)

Add columns:

```{r}
input_data %<>%
  mutate(dwc_kingdom = "Animalia") 
```

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:

```{r}
input_data %<>%
  mutate(dwc_scientificName = `scientific name`) 
```


```{r}
parsed_names <- input_data %>%
  distinct(dwc_scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

Correct names and reparse:

##all names are correct

```{r correct and reparse, eval = FALSE}
#input_data %<>% mutate(scientificName = recode(verbatimScientificName,
#  "Gibbulinella aff. dealbata" = "Gibbulinella dealbata",
#  "Hemicycla cf. gaudryi" = "Hemicycla gaudryi",
#  "Xerotricha aff. orbignii" = "Xerotricha orbignii",
#  "Napaeus cf. venegueraensis" = "Napaeus venegueraensis",
#  "Monilearia cf. praeposita" = "Monilearia praeposita",
#  "Gibbulinella aff. dewinteri" = "Gibbulinella dewinteri",
#  "Pomatias aff. laevigatus" = "Pomatias laevigatus",
#  "Physella cf. acuta" = "Physella acuta",
#  "Hemicycla psathyra cf. temperata" = "Hemicycla psathyra temperata"
#))

# Redo parsing
parsed_names <- input_data %>%
  distinct(scientific_name) %>%
  pull() %>%
  parsenames()

# Show names with nomenclatural issues again
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

## Taxon ranks (not applicable)

The nameparser function also provides information about the rank of the taxon (in `rankmarker`). Here we join this information with our checklist. Cleaning these ranks will done in the Taxon Core mapping:

```{r}
input_data %<>% left_join(
  parsed_names %>%
  select(scientificname, rankmarker),
  by = c("dwc_scientificName" = "scientificname"))
```

```{r}
# Adjust rankmarker
#input_data %<>% mutate(rankmarker = recode(verbatimScientificName,
#  "Monilearia spec." = "genus",
#  "Hemicycla spec." = "genus",
#  "Napaeus spec." = "genus",
#  .default = rankmarker
##))
```

## Taxon IDs (not applicable)



## Preview data

Show the number of taxa and distributions per kingdom and rank:

```{r}
#input_data %>%
#  group_by(kingdom, rankmarker) %>%
#  summarize(
#    `# taxa` = n_distinct(taxon_id),
#    `# distributions` = n()
#  ) %>%
#  adorn_totals("row")
```

Preview data:

```{r}
input_data %>% head()
```

# Darwin Core mapping

## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "PPNEA")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "placeholder") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "PPNEA")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Name to be decided")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

The following terms contain information about the taxon:

### occurrenceID

```{r}
occurrence %<>% mutate(id = row_number())
```


```{r}
occurrence %<>% mutate(dwc_occurrenceID = paste0("PPNEA:BIRDS:OBS:", id))
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = dwc_rightsHolder)
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = number)
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = format(as.Date(date, format="%m/%d/%Y"),"%Y/%m/%d"))
  
```



### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "casual observation")
```

### samplingEffort
 
```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_samplingEffort = case_when(
  !is.na(Mandagen) ~ paste("daysOfWork:", round(Mandagen, digits = 2)),
  !is.na(Val_aantal) ~ paste("numberOfTraps:", floor(Val_aantal)),
  TRUE ~ ""
))
```

### locationID


### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "AL")
```

### stateProvince


### Municipality


### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = lat) 
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = lng) 
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### georeferenceRemarks

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_georeferenceRemarks = case_when(
  endsWith(Coor_prox, "Yes") ~ "coordinates are centroid of municipality",
  endsWith(Coor_prox, "No") ~ "exact coordinates",
  TRUE ~ NA_character_
))
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### scientificName


### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### scientificNameAuthorship


### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "albania", "data", "processed", "occurrence.csv"), na = "")

```
